Spfy: an integrated graph database for real-time prediction of Escherichia coli phenotypes and downstream comparative analyses

### 2018-04-19 paper-webserver.tex
final edits and suggestions

To solve this problem, we created Spfy, which uses a graph database to store and retrieve results from computational workflows, and link data to individual genomes using standardized ontologies.

The Spfy platform facilitates rapid phenotype identification, as well as the efficient storage and downstream comparative analysis of tens of thousands of genome sequences. Though generally applicable to bacterial genome sequences, Spfy currently contains 10,243 \textit{Escherichia coli} genomes, for which \textit{in-silico} serotype and Shiga-toxin subtype, as well as the presence of known virulence factors and antimicrobial resistance determinants have been computed. Additionally, the presence / absence of then entire \textit{E. coli} pan-genome was computed and linked to each genome.

@@

% 2. Big Problem: but tools are for individual analysis
The rapid characterization based on WGS relies on the outputs from multiple software programs that are targeted for specific applications. Examples include the identification of known antimicrobial resistance (AMR) genes, through software such as the Resistance Gene Identifier (RGI) \cite{mcarthur2013comprehensive}, \cite{kleinheinz2014applying}, \cite{gupta2014arg}, and ARIBA \cite{hunt2017ariba}; or the identification of known virulence factor genes (VF) through software such as VirulenceFinder \cite{kleinheinz2014applying}, SRST2 \cite{inouye2014srst2}, and GeneSippr \cite{lambert2015genesippr}. For clinical diagnoses and comparisons, individual species can be first divided into subtypes with complementing AMR and VF results. Software methods for subtyping rely on pre-selected intraspecies genes or genomic regions, and are targeted through software such as Phylotyper \cite{whiteside2017phylotyper}, SerotypeFinder \cite{joensen2015rapid}, the EcOH dataset applied through SRST2 \cite{ingle2016silico}, and V-Typer \cite{carrillo2016comparative}. These methods represent \textit{in-silico} analogues of traditional wet-lab tests, which allows new WGS results to be viewed in the context of historical tests, and greatly expedites the analyses of newly sequenced genomes.

@@

Many of the analyses that are currently used in the characterization and study of bacterial genomes are broadly useful, and therefore the same analyses are often computed multiple times for the same genome. An effective method to mitigate this re-computation, is to make the storage and retrieval of results part of the analyses platform, and effectively linked to the genomes of interest through a standardized ontology. Such measures help ensure the rapid response times required for public health applications, and allow results to be integrated and progressively updated as new data becomes available.

@@

% 5. how we solved it
Here, we present a new platform merging Superphy's pre-computed results with a novel data storage and processing architecture, which we call Spfy; Spfy integrates a graph database with real-time analyses to avoid recomputing identical results. Graph-based result storage also allows retrospective comparisons across stored results as more genomes are sequenced or populations change, and is flexible, accommodating new analysis modules as they are developed. The database is available at \url{https://lfz.corefacility.ca/superphy/spfy/}.

@@

Spfy handles all of the analyses tasks by dividing them into subtasks, which are subsequently distributed across a built-in task queue. Results are converted into individual graphs and stored within a larger graph database according to the previously created ontologies GenEpiO \cite{griffiths2017context}, FALDO \cite{bolleman2016faldo}, and TypOn \cite{vaz2014typon}, where metadata including genotypes, biomarkers, host, source are stored.

By integrating task distribution with graph storage, Spfy enables large-scale analyses, such as epidemiological association studies. Any data-type or relation in the graph is a valid option for analysis. This means that genomes can be compared on the basis of the presence or absence of pan-genome regions, serotype, subtyping data, or provided metadata such as location or host-source. All results are displayed to users in real-time, usually within 2-3 minutes.
For example, Spfy can determine if there's a statistically significant difference, using Fisher's exact test, in the presence of all known AMR genes in collected O157 \textit{E. coli} genomes vs. O26 as shown in Figure \ref{fig-groupings}.
Different types can also be joined into a group through logical connectives AND, OR, or Negation. This approach can be used to compare any data regardless of source software.

\begin{figure}[!hb]
\begin{center}
\includegraphics[width=\textwidth]{images/o157vo26_amrgenes}
\end{center}
\caption{
% This needs to be specific to the image we are showing eg. the O157 vs. O26 comparison. We mention the general comparison in the text, and this can be offered as a specific example of that.
Genomes with the predicted subtype O157 are compared against the subtype O26. In this example, there are 903 O157 genomes and 291 O26 genomes which are compared for the presence of 129 AMR genes. Fisher's Exact Test is used for the comparison.}
\label{fig-groupings}
\end{figure}

\section{IMPLEMENTATION}
% para: intro to the spfy stack
The server-side code for Spfy, graph generation, and analysis modules, are developed in Python, with the front-end website developed using the React JavaScript library \url{https://facebook.github.io/react/}. When new data is added to the database, the following steps are taken:

@@

\subsection{Data Storage}
% para
% 0. Goals: big-data, everything linked, easy addition of new links
% 1. spfy is built around graph technologies
% 1. how we structure our graph
% 2. ontoogies used
% 3. inferencing
% 3 1/2. SPARQL queries
% para: the semantic web
Semantic web technologies describe the relationships between data, and have been proposed as an open standard for sharing public information \cite{berners2001semantic}, while graph databases are a flexible means of storing this this information \cite{horrocks2005semantic}. Biological data can be a genome, contiguous DNA sequence, or gene, and these are linked together in a searchable graph structure using existing ontologies. This system is flexible and allows novel data to be incorporated into the existing graph.

% The text in this figure is too small to read and seems blurry. It should be enlarged.
\begin{figure}[!hb]
\begin{center}
\includegraphics[width=\textwidth]{images/ontology}
\end{center}
\caption{Structure of the Spfy graph database. Brackets highlight the source of different data points and the software it was generated from. Data are added in as the analysis modules complete, at varying times, and the overall connections are inferred by the database. Non-bracketed sections are sourced from the uploaded genome files or user-supplied metadata.}
\label{fig-ontology}
\end{figure}

@@

\section{RESULTS}
Spfy was tested with 10,243 public \textit{E. coli} assembled genomes from Enterobase, storing every sequence and the results for all included analysis modules. This included: serotyping (O-antigen, H-antigen), toxin sub-typing (Shiga-toxin 1, Shiga-toxin 2, and Intimin), the identification of VF and AMR determinants, and determination of the pan-genome content of \textit{E. coli}.
The resulting database has 17,820 nodes and 3,811,473 leaves, with 1,125,909,074 object properties.
Spfy has been up since May 2017. The server accepts assembled \textit{E. coli} genomes with the \textit{.fasta} or \textit{.fna} extensions. Submissions are subjected to quality control, ensuring the submitted genomes are \textit{E. coli} sequences before subsequent analyses are run.
\par

@@

Spfy is running on a virtual machine (VM) with 8 vCPUs emulating single-core Intel® Xeon® E3-series processors, and with 32 GB of RAM.
The VM is running CentOS version 7.3, with Docker version 17.06.1-ce and Docker-Compose version 1.12.0.
On comparison tasks, Spfy can retrieve and compare 1 million nodes/attributes in the graph database in aproximately 70 seconds.
As shown in Figure \ref{fig:fishers_performance}, performance scales linearly with the number of genomes involved in the comparison, or the number of target nodes retrieved per genome; 1.5 million nodes/attributes can be compared in aproximately 90 seconds, and 2 million nodes/attributes in aproximately 110 seconds.

\begin{figure}[!htb]
\minipage{0.32\textwidth}
  \includegraphics[width=\linewidth]{images/fishers_genomes.png}
\endminipage\hfill
\minipage{0.32\textwidth}
  \includegraphics[width=\linewidth]{images/fishers_targets.png}
\endminipage\hfill
\minipage{0.32\textwidth}%
  \includegraphics[width=\linewidth]{images/fishers_overall.png}
\endminipage
\caption{Runtimes of Fisher's Exact Test depending on the number of nodes/attributes involved in the comparison. Left: Runtimes as the number of genomes involved increases for a fixed (107) number of targets per genome. Center: Runtimes as the number of targets involved increased for a fixed (116) number of genomes per target. Right: Runtimes as the total number of targets retrieved increases; the total number of targets is calculated as follows: (Number Genomes Group A + Number Genomes Group B) x Number Targets per Genome. In all cases, we see a linear increase in runtime as the number of targets or the number of genomes involved increases.}\label{fig:fishers_performance}
\end{figure}

On analysis tasks, Spfy runs all included modules in an average of 130 seconds per genome, as shown in Figure \ref{fig:spfy_performance_indivs}.
Spfy can also queue batches of genomes for analysis, decreasing the average runtime to an average of 54 seconds per file due to parallelization of analysis runs \ref{fig:spfy_performance_batches}.
In total, the platform analyzed 50 genomes in 45 minutes, and 100 genomes in 89 minutes.

\begin{figure}[!hb]
\begin{center}
\includegraphics[width=\textwidth]{images/spfy_indivs.png}
\end{center}
\caption{Runtimes of Spfy's analysis modules. 100 genome files were run one-by-one with the different colors representing the runtimes of the different analysis module; The blue line indicates the average actual runtime per file after accounting for parallelization.}\label{fig:spfy_performance_indivs}
\end{figure}

\begin{figure}[!htb]
\minipage{0.5\textwidth}
  \includegraphics[width=\linewidth]{images/spfy_batches_totals.png}
\endminipage\hfill
\minipage{0.5\textwidth}%
  \includegraphics[width=\linewidth]{images/spfy_batches.png}
\endminipage
\caption{Runtimes of Spfy's analysis modules for batches of files. Left: Total runtime for analyzing all files in the given batch. The blue line indicates the actual time to completion after accounting for parallelization; 50 files are analyzed in 45 minutes and 100 files in 89 minutes. Right: Average runtime per file in the batch. As indicated by the line, the average runtime per genome in consistant when running more than one file. This is due to the degree of parallelization reaching the capacity of the host computer.}\label{fig:spfy_performance_batches}
\end{figure}

@@

% This has not been shown in this paper; it has only been stated. We should probably have a large-scale analysis for an example, with accompanying times etc.
We have shown how a graph database can accommodate the results from a variety of bioinformatics programs, and how Spfy is performant for data retrieval on the results for multiple analyses among over 10,000 genomes, providing results from big-data comparisons in the same efficiency as old analyses on single files.

@@

% namely galaxy
The automated analyses of WGS is currently facilitated by existing scientific workflow technologies such as Galaxy \cite{goecks2010galaxy}. Galaxy aims to provide a reproducible, computational interface which is accessible to individuals without programming knowledge. Galaxy defines a formal schema for linking different analysis software together, so the entire pipeline can be replicated and also extended as new tools are developed. The Galaxy workflow focuses on running an individual analysis pipeline. It does not include functionality to store and collate analysis results for large-scale comparative studies.

@@

Spfy is similar to these technologies in that it automates workflows for users and uses task queues to distribute selected analyses.
% re: Matt "2. When comparing to IRIDA, BAP etc., can mention some differences with Superphy, e.g. the storage of interim result data that allows downstream integrated analysis"
% If we mention speed comparisons, we need to have actual data supporting them
On a per file basis, Spfy performs at a similar speed to BAP on predictive genomics tasks, though Spfy does not provide genome assembly services.
After accounting for assembly services, BAP reported \cite{thomsen2016bacterial} an average runtime of 8-9 minutes per genome over 476 runs which is in the same scale as Spfy's average of over 2 minutes.
Note that the two platforms are now directly comparable, due to the differences in analysis tasks invovled.
In similar tasks such as AMR determination, the ResFinder program included in BAP took an average of 3-4 minutes \cite{thomsen2016bacterial} and is similar to the RGI program included in Spfy, which took an average of 1 minute 30 seconds.

@@

PATRIC \cite{wattam2016improvements} and Spfy share the same goal of integrated analyses. PATRIC has support for comparing up to nine user-submitted genomes against a reference genome, based on gene annotations; the platform indexes a NoSQL document store to compare similar document types. Unlike PARTIC, Spfy provides the ability to perform statistical comparisons of any permutation of a population group based on the chosen data types, and Spfy's graph database has no limit on the number of genomes grouped for comparison.

@@

The integrated approach taken in the creation of Spfy, where the storage and retrieval of results provides enormous benefits for the large-scale analyes of \textit{E. coli}. The developed analyses modules are also self-contained and be transferred to existing platforms such as Galaxy. Future work will focus on adding machine learning modules to improve genotype / phenotype predictions, and supporting bacterial species such as \textit{Salmonella}, and \textit{Campylobacter}. The source code for Spfy is hosted at \url{https://github.com/superphy/backend}, and is available for free under the open-source Apache 2.0 license. A developer guide is provided at \url{https://superphy.readthedocs.io/en/latest/}.

>>>>

To solve this problem, we created Spfy, a platform that rapidly perform the common reference lab tests, uses a graph database to store and retrieve the results from the computational workflows, and links data to individual genomes using standardized ontologies.

The Spfy platform facilitates rapid phenotype identification, as well as the efficient storage and downstream comparative analysis of tens of thousands of genome sequences. Though generally applicable to bacterial genome sequences, Spfy currently contains 10,243 \textit{Escherichia coli} genomes, for which \textit{in-silico} serotype and Shiga-toxin subtype, as well as the presence of known virulence factors and antimicrobial resistance determinants have been computed. Additionally, the presence / absence of the entire \textit{E. coli} pan-genome was computed and linked to each genome.

@@

% 2. Big Problem: but tools are for individual analysis
The rapid characterization based on WGS relies on the outputs from multiple software programs that are targeted for specific applications. Examples include the identification of known antimicrobial resistance (AMR) genes, through software such as the Resistance Gene Identifier (RGI) \cite{mcarthur2013comprehensive}, ResFinder \cite{kleinheinz2014applying}, ARG-ANNOT \cite{gupta2014arg}, and ARIBA \cite{hunt2017ariba}; or the identification of known virulence factor genes (VF) through software such as VirulenceFinder \cite{kleinheinz2014applying}, SRST2 \cite{inouye2014srst2}, and GeneSippr \cite{lambert2015genesippr}.

For subtyping in clinical diagnoses and reference lab environments, software relies on pre-selected intraspecies genes or genomic regions, which are targeted through software such as Phylotyper \cite{whiteside2017phylotyper}, SerotypeFinder \cite{joensen2015rapid}, the EcOH dataset applied through SRST2 \cite{ingle2016silico}, and V-Typer \cite{carrillo2016comparative}. These methods represent \textit{in-silico} analogues of traditional wet-lab tests, which allow new whole-genome sequences to be viewed in the context of historical tests without the need for the time and labor of the traditional wet-lab tests.

@@

Many of the analyses used in the characterization and study of bacterial genomes are broadly useful, and therefore they are often computed multiple times for the same genome across different studies. An effective method to mitigate this re-computation, is to make the storage and retrieval of results part of the analyses platform, and effectively linked to the genomes of interest through a standardized ontology. Such measures help ensure the rapid response times required for public health applications, and allow results to be integrated and progressively updated as new data becomes available.

@@

% 5. how we solved it
Here, we present a new platform merging the pre-computed results of SuperPhy, with a novel data storage and processing architecture, which we call Spfy; Spfy integrates a graph database with real-time analyses to avoid recomputing identical results. Additionally, graph-based result storage allows retrospective comparisons across stored results as more genomes are sequenced or populations change. Spfy is flexible, accommodating new analysis modules as they are developed. The database is available at \url{https://lfz.corefacility.ca/superphy/spfy/}.

@@

Spfy handles all of the analyses tasks by dividing them into subtasks, which are subsequently distributed across a built-in task queue. Results are converted into individual graphs and stored within a larger graph database according to the previously created ontologies GenEpiO \cite{griffiths2017context}, FALDO \cite{bolleman2016faldo}, and TypOn \cite{vaz2014typon}, which includes metadata for genotypes, location, biomarkers, host, and source, among others.

By integrating task distribution with graph storage, Spfy enables large-scale analyses, such as epidemiological association studies. Any data-type or relation in the graph is a valid option for analysis. This means that genomes can be compared on the basis of the presence or absence of pan-genome regions, serotype, subtyping data, or provided metadata such as location or host-source. All results are displayed to users in real-time, usually within 2-3 minutes.
For example, Spfy can determine if a statistically significant difference exists, using Fisher's exact test, among any identified AMR genes, between \textit{E. coli} genomes of serotype O157, and genomes of serotype O26, as shown in Figure \ref{fig-groupings}.
Different types can also be joined into a group through logical connectives AND, OR, or Negation. This approach can be used to compare any data regardless of source software.

\begin{figure}[!hb]
\begin{center}
\includegraphics[width=\textwidth]{images/o157vo26_amrgenes}
\end{center}
\caption{
% This needs to be specific to the image we are showing eg. the O157 vs. O26 comparison. We mention the general comparison in the text, and this can be offered as a specific example of that.
\textit{E. coli} genomes of serotype O157 (903) compared against genome of serotype O26 (291), for statistically significant differences in the carriage of 129 AMR genes. Fisher's Exact Test is used by Spfy for these comparisons.}
\label{fig-groupings}
\end{figure}

\section{IMPLEMENTATION}
% para: intro to the spfy stack
The server-side code for Spfy, graph generation, and analysis modules, were developed in Python, with the front-end website developed using the React JavaScript library \url{https://facebook.github.io/react/}. When new data is added to the database, the following steps are taken:

@@

\subsection{Data Storage}
% para
% 0. Goals: big-data, everything linked, easy addition of new links
% 1. spfy is built around graph technologies
% 1. how we structure our graph
% 2. ontoogies used
% 3. inferencing
% 3 1/2. SPARQL queries
% para: the semantic web
Semantic web technologies describe the relationships between data, and have been proposed as an open standard for sharing public information \cite{berners2001semantic}, while graph databases are a flexible means of storing this information \cite{horrocks2005semantic}. Biological data can be a genome, contiguous DNA sequence, or gene, and these are linked together in a searchable graph structure using existing ontologies. This system is flexible and allows novel data to be incorporated into the existing graph.

% This is still true. The text needs to be larger: --> The text in this figure is too small to read and seems blurry. It should be enlarged.
\begin{figure}[!hb]
\begin{center}
\includegraphics[width=\textwidth]{images/ontology}
\end{center}
\caption{Structure of the Spfy graph database. Brackets highlight the source of different data points and the software it was generated from. Data are added in as the analysis modules complete, at varying times, and the overall connections are inferred by the database. Non-bracketed sections are sourced from the uploaded genome files or user-supplied metadata.}
\label{fig-ontology}
\end{figure}

@@

\section{RESULTS}
Spfy was tested with 10,243 public \textit{E. coli} assembled genomes from Enterobase, storing every sequence and the results for all included analysis modules. These included: serotyping (O-antigen, H-antigen), toxin sub-typing (Shiga-toxin 1, Shiga-toxin 2, and Intimin), the identification of VF and AMR determinants, and determination of the pan-genome content of \textit{E. coli}.
The resulting database has 17,820 nodes and 3,811,473 leaves, with 1,125,909,074 object properties.
Spfy has been up since May 2017. The server accepts assembled \textit{E. coli} genomes with the \textit{.fasta} or \textit{.fna} extensions. Submissions are subjected to quality control, ensuring the submitted genomes are \textit{E. coli} sequences before subsequent analyses are run.
\par

@@

Spfy is running on a virtual machine (VM) with 8 vCPUs emulating single-core Intel® Xeon® E3-series processors, and with 32 GB of RAM.
The VM is running CentOS version 7.3, with Docker version 17.06.1-ce and Docker-Compose version 1.12.0.
On comparison tasks, Spfy can retrieve and compare 1 million nodes/attributes in the graph database in approximately 70 seconds.
As shown in Figure \ref{fig:fishers_performance}, performance scales linearly with the number of genomes involved in the comparison, or the number of target nodes retrieved per genome; 1.5 million nodes/attributes can be compared in approximately 90 seconds, and 2 million nodes/attributes in approximately 110 seconds.

% The X and Y axes labels need to be slightly larger. Designate the panes a), b), and c) and then reference them in the caption. The caption should contain the information that the titles on each pane currently do.
\begin{figure}[!htb]
\minipage{0.32\textwidth}
  \includegraphics[width=\linewidth]{images/fishers_genomes.png}
\endminipage\hfill
\minipage{0.32\textwidth}
  \includegraphics[width=\linewidth]{images/fishers_targets.png}
\endminipage\hfill
\minipage{0.32\textwidth}%
  \includegraphics[width=\linewidth]{images/fishers_overall.png}
\endminipage
\caption{Runtimes of Fisher's Exact Test depending on the number of nodes/attributes involved in the comparison. Left: Runtimes as the number of genomes increased for a fixed (107) number of targets per genome. Center: Runtimes as the number of targets increased for a fixed (116) number of genomes per target. Right: Runtimes as the total number of targets retrieved increased; the total number of targets was calculated as follows: (Number Genomes Group A + Number Genomes Group B) x Number Targets per Genome. In all cases, a linear increase in runtime was observered as the number of targets or genomes increased.}\label{fig:fishers_performance}
\end{figure}

On analysis tasks, Spfy runs all included analyses modules in an average of 130 seconds per genome, as shown in Figure \ref{fig:spfy_performance_indivs}.
Spfy can also queue batches of genomes for analysis, decreasing the average runtime to an average of 54 seconds per file due to parallelization of analysis runs \ref{fig:spfy_performance_batches}.
In total, the platform analyzed 50 genomes in 45 minutes, and 100 genomes in 89 minutes.

\begin{figure}[!hb]
\begin{center}
\includegraphics[width=\textwidth]{images/spfy_indivs.png}
\end{center}
\caption{Runtimes of Spfy's analysis modules. 100 genomes were run individually, with each coloured dot representing the runtime of a separate analysis module. The blue line indicates the average actual runtime per file after accounting for parallelization.}\label{fig:spfy_performance_indivs}
\end{figure}

% Again, label panes as a) and b), then reference in the caption
\begin{figure}[!htb]
\minipage{0.5\textwidth}
  \includegraphics[width=\linewidth]{images/spfy_batches_totals.png}
\endminipage\hfill
\minipage{0.5\textwidth}%
  \includegraphics[width=\linewidth]{images/spfy_batches.png}
\endminipage
\caption{Runtimes of Spfy's analysis modules for batches of files. Left: Total runtime for analyzing all files in the given batch. The blue line indicates the actual time to completion after accounting for parallelization; 50 files are analyzed in 45 minutes and 100 files in 89 minutes. Right: Average runtime per file in the batch. As indicated by the line, the average runtime per genome in consistent when running more than one file due to the parallelization of analyses.}\label{fig:spfy_performance_batches}
\end{figure}

@@

% This has not been shown in this paper; it has only been stated. We should probably have a large-scale analysis for an example, with accompanying times etc.
We have shown how a graph database can accommodate the results from a variety of bioinformatics programs, and how Spfy is performant for data retrieval of the results from multiple analyses among over 10,000 genomes. Spfy provides results from big-data comparisons with the same efficiency as old analyses on single files.

@@

% namely galaxy
The automated analyses of WGS is currently facilitated by existing scientific workflow technologies such as Galaxy \cite{goecks2010galaxy}. Galaxy aims to provide a reproducible, computational interface that is accessible to individuals without programming knowledge. Galaxy defines a formal schema for linking different analysis software together, so the entire pipeline can be replicated and also extended as new tools are developed. The Galaxy workflow focuses on running an individual analysis pipeline. It does not include functionality to store and collate analysis results for large-scale comparative studies.

@@

Spfy is similar to these technologies in that it automates workflows for users and uses task queues to distribute selected analyses.
% re: Matt "2. When comparing to IRIDA, BAP etc., can mention some differences with Superphy, e.g. the storage of interim result data that allows downstream integrated analysis"
% If we mention speed comparisons, we need to have actual data supporting them
On a per file basis, Spfy performs at a similar speed to BAP on predictive genomics tasks, though Spfy does not provide genome assembly services.
After accounting for assembly services, BAP reported \cite{thomsen2016bacterial} an average runtime of 8-9 minutes per genome over 476 runs which is in the same scale as Spfy's average of over 2 minutes.
Note that the two platforms are not directly comparable, due to the differences in analysis tasks involved.
In similar tasks such as AMR determination, the ResFinder program included in BAP took an average of 3-4 minutes \cite{thomsen2016bacterial} and is similar to the RGI program included in Spfy, which took an average of 1 minute 30 seconds.

@@

PATRIC \cite{wattam2016improvements} and Spfy share the same goal of integrated analyses. PATRIC has support for comparing up to nine user-submitted genomes against a reference genome, based on gene annotations; the platform indexes a NoSQL document store to compare similar document types. Unlike PARTIC, Spfy provides the ability to perform statistical comparisons of any permutation of a population group based on the chosen data types, and the graph database of Spfy has no limit on the number of genomes grouped for comparison.

@@

The integrated approach taken in the creation of Spfy, where the analyses, storage, and retrieval of results is combined, provides enormous benefits for the large-scale analyses of \textit{E. coli}. The developed analyses modules are also self-contained and can be used in existing platforms such as Galaxy. Future work will focus on adding machine learning modules to improve genotype / phenotype predictions, and supporting bacterial species such as \textit{Salmonella}, and \textit{Campylobacter}. The source code for Spfy is hosted at \url{https://github.com/superphy/backend}, and is available for free under the open-source Apache 2.0 license. A developer guide is provided at \url{https://superphy.readthedocs.io/en/latest/}.
